---
title: "Final Project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#I-) Essential Packages
```{r}
IscaretInstalled <- require("caret")
if(!IscaretInstalled){
    install.packages("caret")
    library("caret")
    }

IsrandomForestInstalled <- require("randomForest")
if(!IsrandomForestInstalled){
    install.packages("randomForest")
    library("randomForest")
    }

IsRpartInstalled <- require("rpart")        
if(!IsRpartInstalled){
    install.packages("rpart")
    library("rpart")
    }

#set seed for reproducability
set.seed(20000)
```

#II-) Data Processing
Load the data
```{r}
trainUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"   
testUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv" 
```

#III-)Cleaning data 
Load data to memory.
```{r}
training <- read.csv(url(trainUrl), na.strings=c("NA","#DIV/0!",""))  
testing <- read.csv(url(testUrl), na.strings=c("NA","#DIV/0!",""))

#Remove variables with near zero variance
training<-training[,colSums(is.na(training)) == 0]
testing <-testing[,colSums(is.na(testing)) == 0]

#Remove columns that are not predictors, which are the the seven first columns
training   <-training[,-c(1:7)]
testing <-testing[,-c(1:7)]

# The data after cleaning
dim(training)
dim(testing)
```
#IV-) Cross-validation
In order to get out-of-sample errors, split the training data in training (75%) and testing (25%) data) subsets:
```{r}
inTrain <- createDataPartition(y=training$classe, p=0.75, list=FALSE)    
NEOTraining <- training[inTrain, ]
NEOTesting <- training[-inTrain, ]  
dim(NEOTraining)
dim(NEOTesting) 
```

#V-) Prediction Models

DECISION TREE
Fit model on NEOTraining data
```{r}
fitDT <- rpart(classe ~ ., data=NEOTraining, method="class")
```
Use model to predict class in validation set (NEOTesting)
```{r}
predictionDT <- predict(fitDT, NEOTesting, type = "class")
```
Estimate the errors of the prediction algorithm in the Decision Tree model
```{r}
NEOTesting$classe <- as.factor(NEOTesting$classe)
confusionMatrix(predictionDT,NEOTesting$classe)
```

RANDOM FOREST
Fit model on NEOTraining data
```{r}
NEOTraining$classe <- as.factor(NEOTraining$classe)
fitRF <- randomForest(classe ~ ., data=NEOTraining, method="class")
```
Use model to predict class in validation set (NEOTesting)
```{r}
predictionRF <- predict(fitRF, NEOTesting, type = "class")
```
Estimate the errors of the prediction algorithm in the Random Forest
```{r}
confusionMatrix(NEOTesting$classe, predictionRF)
```

#VI-) TEST THE MODEL TO PREDICT 20 DIFFERENT TEST CASES
# Perform prediction
```{r}
predictSubmission <- predict(fitRF, testing, type="class")
predictSubmission
```
